Scaling rates of true polar wander in convecting planets and moons

### 2016-08-06 tpw_rate.tex
Fix reference.

Let $\delta$ be the size of the perturbation to the convective moment of inertia tensor after some time interval $\Delta t$, and let $\lambda_3 \ge \lambda_2 \ge \lambda_1$ be the eigenvalues of that tensor.  The corresponding rotation of the principal axes is defined by the angle $\xi$. A bound on $\xi$ is given by 
\begin{equation}
\vert \sin(2 \xi) \vert \le \frac{ 2 \vert \delta \vert}{ \displaystyle \min_{i \neq j} \vert \lambda_i - \lambda_j \vert }.
\label{eq:kahan}
\end{equation} 
That is to say, if there is a large difference between the eigenvalues, this stabilizes axes to perturbations.  
If, however, there is a small difference between the eigenvalues (i.e., they are nearly degenerate), then perturbations can cause large rotations of the principal axes.
This is illustrated schematically in Figure~\ref{fig:perturb}.
\citet{evans1998} described a situation where the convective moment of inertia of Earth is prolate, with the
maximum and intermediate moments of inertia nearly equal.
He argued that perturbations in this case would result in TPW paths that are approximately
on the great circle path between the intermediate and maximum axes.
This behavior is a consequence of the instability described in Equation~\eqref{eq:kahan}.

>>>>

Let $\delta$ be the size of the perturbation to the convective moment of inertia tensor after some time interval $\Delta t$, and let $\lambda_3 \ge \lambda_2 \ge \lambda_1$ be the eigenvalues of that tensor.  The corresponding rotation of the principal axes is defined by the angle $\xi$. A bound on $\xi$ is given by 
\begin{equation}
\vert \sin(2 \xi) \vert \le \frac{ 2 \vert \delta \vert}{ \displaystyle \min_{i \neq j} \vert \lambda_i - \lambda_j \vert }.
\label{eq:kahan}
\end{equation} 
That is to say, if there is a large difference between the eigenvalues, this stabilizes axes to perturbations.  
If, however, there is a small difference between the eigenvalues (i.e., they are nearly degenerate), then perturbations can cause large rotations of the principal axes.
This is illustrated schematically in Figure~\ref{fig:perturb}.
\citet{evans1998true} described a situation where the convective moment of inertia of Earth is prolate, with the
maximum and intermediate moments of inertia nearly equal.
He argued that perturbations in this case would result in TPW paths that are approximately
on the great circle path between the intermediate and maximum axes.
This behavior is a consequence of the instability described in Equation~\eqref{eq:kahan}.
